---
title: "Assignment Template"
output:
  html_document: default
  pdf_document: default
date: "2024-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*The following is a template .rmd RMarkdown file for you to use for your homework submission.*

*Please Knit your .rmd to a PDF format or HTML and submit that with no identifiers like your name.*

*To create a PDF, first install tinytex and load the package. Then press the Knit arrow and select "Knit to PDF".*

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r bad figure code, echo=FALSE}
echo = FALSE
library(tidyverse)
library(palmerpenguins)
library(here)
library(janitor)
library(ggplot2)
library(ggbeeswarm)
library(ragg)
library(svglite)
library(patchwork)
library(scatterplot3d)
library(dplyr)
library(car)

source(here("functions", "cleaning.R"))

write_csv(penguins_raw, here("data", "penguins_raw.csv"))

penguins_clean <- penguins_raw %>%
  clean_column_names() %>%
  remove_columns(c("comments", "delta")) %>%
  shorten_species() %>%
  remove_empty_columns_rows() %>%
  remove_NA()

sex_numeric <- as.numeric(factor(penguins_clean$sex)) - 1

badplot <- scatterplot3d(x = penguins_clean$flipper_length_mm,
                     y = penguins_clean$body_mass_g,
                     z = sex_numeric,
                     color = rainbow(nrow(penguins_clean)),
                     main = "Bad Penguin Plot",
                     xlab = "flipper Length ",
                     ylab = "bodymass ",
                     zlab = "Sex ",
                     angle = 70,
                     scale.y = 0.7,
                     box = FALSE)
                     
```


### b) Write about how your design choices mislead the reader about the underlying data (100-300 words).

This plot has many problems that prevent us from being able to correctly or easily interpret the data or any trends that exist. These issues range from aesthetically irritating characteristics (such as inconsistent capitalisation and spacing on the graph axes), to more serious problems such as arbitrary axis start points, and a confusing 3D structure. While there is nothing that is explicitly a ‘lie’ (i.e. the data points have not been changed, and the variable names still align with those in the table), the graph in its current form is next to meaningless. The fact that ‘Sex’ is represented as a numerical value (despite only 1 and 0 being available as ‘male’ and ‘female’), is problematic, as is the fact that it is not made clear which is which. Additionally, there are no units of measurement for any of the axes, which makes it difficult to draw meaningful conclusions from the data presented. The fact that the flipper length and body mass axes don’t start at 0 is also a problem, as this can over represent any difference found between different data points (i.e. there appears to be a greater distance between 170 and 200 on a graph if the axes start at 170, than if they start at 0). The use of a rainbow gradient of plot points is also pointless, and doesn’t use the opportunity to use colour to add additional information in a meaningful way (for example, if there were 3 different colours for points, each representing a different species, this would at least give some indication of how flipper length might vary between species and sex. Finally, the 3D nature of the plot (without any background lines), makes the body mass values impossible to interpret.

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps as well as clear code.*

*Your code should include the steps practiced in the lab session:*

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

-   ***New**: Run a statistical test*

-   ***New**: Create a Results Figure*

-   *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test or model.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

### Introduction

These are the packages that will be used. They need to be installed, and loaded using the library() function. We can then take a quick look at the dataset using View() or glimpse()
```{r Data Exploration}

# Make sure your code prints. 
#renv::snapshot()

library(tidyverse)
library(palmerpenguins)
library(here)
library(janitor)
library(ggplot2)
library(ggbeeswarm)
library(ragg)
library(svglite)
library(patchwork)
library(car)
library(dplyr)


```
As the column names are inconsistent, and there are gaps in the data, it requires cleaning.

```{r}
penguins_clean <- penguins_raw %>% 
  clean_column_names() %>% # this first 'task', using the clean_names() function from the janitor package to increase the uniformity of the column names and to remove any characters that may cause issues in commands further down the line
  remove_columns(c("comments", "delta")) %>% # this line removes the comments an delta columns. The 'Comments' contains qualitative data and descriptions that may be useful in other contexts, but they cannot be used for statistical tests or graphical analysis. Likewise, the  'Delta' columns contain isotope values that are not going to be used in the following data analyses.
  shorten_species() %>% # this function reduces the length of the species names for ease of use in data processing downstream
  remove_empty_columns_rows() %>% # this removes any columns or rows that are entirely empty or contain only NA
  remove_NA() # this will remove any data points that have NA values, which can cause problems for calculations and graphical representations later on

write_csv(penguins_clean, here("data", "penguins_clean.csv")) 
```

```{r}
#this sets particular colours for each species so that they can be visualised
species_colours <- c("Adelie" = "darkgrey", "Chinstrap" = "blue", "Gentoo" = "cyan4")

# This creates an exploratory graph (a beeswarm plot in this case)
ggplot(data = penguins_clean, aes(x = sex, y = body_mass_g, colour = species)) +
  geom_beeswarm(cex = 2, size = 1.5) +
  scale_color_manual(values = species_colours) +
  labs(title = "Body Mass Distribution by Sex in Penguins",
       x = "Sex",
       y = "Body Mass (g)",
       colour = "Species") +
  scale_y_continuous(limits = c(0, 8000), breaks = seq(0, 8000, by = 1000)) +  # this makes the Y axis go up in 1000g intervals
  theme_minimal() # this removes the grey background from the graph
```

### Hypothesis

There appears to be an interesting distribution for body mass, so we could test if there is a stronger relationship between sex and body mass, or species and body mass, as well as any interaction effects. Therefore I will carry out a two-way ANOVA test.
Just from looking at the exploratory plot, I would predict that male penguins (on average) are heavier than female penguins, and that there is a significant difference between the weight of Gentoo penguins in comparison to Chinstrap or Adelie penguins.
To check that the assumptions of the ANOVA are met, the Shapiro-Wilks test and Levene test also need to be carried out.

### Statistical Methods
I will carry out a two way ANOVA test.
```{r Statistics}

#this fits a model to the data
penguinModel <- lm(body_mass_g ~ sex * species, data = penguins_clean)

#these aid in testing the assumption of normality
qqnorm(residuals(penguinModel))
qqline(residuals(penguinModel))

shapiro.test(residuals(penguinModel))

#this tests the homogeneity of variance
leveneTest(body_mass_g ~ sex * species, data = penguins_clean)

#the Shapiro Wilks test  comes back with a p value of 0.9367, indicating normality
#The Levene's test comes back with a p value of 0.2272, indicating homogeneity of variance

#Statistical Analysis of the model:
Anova(penguinModel)
summary(penguinModel)
```

### Results & Discussion

```{r Plotting Results}

#graph indicating the presence of interactions between sex and species on body mass
interaction_plot <- ggplot(penguins_clean, aes(x = species, y = body_mass_g, color = sex, group = sex)) +
  stat_summary(fun = mean, geom = "point", size = 4) + # this adds the means
  stat_summary(fun = mean, geom = "line") + # this adds lines (parallel indicates no interaction, anything else indicates an interaction effect)
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) + # This adds error bars
  scale_y_continuous(limits = c(0, 6000)) + #this sets the limits of the y axis between 0 and 6000
  labs(
    title = "Interaction of Sex and Species on Penguin Body Mass",
    x = "Species",
    y = "Body Mass (g)",
    color = "Sex",
    linetype = "Sex"
  ) +
  theme_minimal()
```

### Conclusion

The results of the ANOVA table suggest that there is a highly significant effect of sex on body mass (p<0.00001), but that the species also greatly impacts body mass (0.00001). There is also an interaction effect, as in Chinstrap penguins, males are slightly lighter than females , but in Gentoo penguins, they are slightly heavier. Overall, if controlling for species, males are heavier than females. In general, Gentoo penguins are heavier than Chinstrap or Adelie penguins. Both of these results align with the predictions made after viewing the exploratory plot.
------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:* https://github.com/zhouyu-jilu/ReproducibleFigures.git

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:* https://github.com/frogdog22/reproducibleFigures.git

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand and run their data pipeline?*
Clear code comments and writing only a few lines per chunk made it easy to see the order of the code and what it did.

-   *Did it run? Did you need to fix anything?*
I think that part of the reason why I had difficulties running it initially was due to problems with PositCloud (I am unable to use the offline version of R as my laptop is currently non functional, the library computers don't allow R to be downloaded, and neither does the college laptop that was loaned to me). This particularly seemed to be the case for anything involving renv. However, if I just installed the packages initially when prompted, they would load without issues.  The first part of the code involving sources (i.e. source(here("functions","libraries.r"))). However, I saw that the files were actually .R not .r, so after changing the code to  source(here("functions","libraries.R")), I was able to run the chunk. On the other hand, for plotting the first figure, I initially received error messages due to the deprecated use of a particular line of code, which a new version of ggplot2 no longer recognised. This may have been avoided if renv had functioned properly. As it hadn't, I searched up the error and altered the function in the function file, to match the updated code requirements. A similar deprecation problem was found in the figure analysis chunk, and was likewise updated (at which point it ran correctly). On the other hand, I was unable to render the HTML from the file, unless I commented out all of the 'save as png' sections (after which it rendered without issue).

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*
In general I found the code well explained and understandable. I still believe that had renv functioned (which my partner had set up, and I am not sure why it did not work), I would not have encountered to deprecated code errors, and as such the code would have run smoothly and achieved its intended purpose. There is nothing in particular that I would suggest would make it easier. Perhaps (as an inexperienced coder) it would have been simpler for me to not use functions that were separated out in another file. However, for someone more familiar with coding, I am sure that separate function files are more uesful.
-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*
I think altering my partner's figures should be easy enough with some basic understanding of R, as they had commented much of what their code did, and also had separated things into the function folder. If necessary, direct edits could be made there.

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*
My partner recommended that I use a separate file for graph making functions rather than putting the code directly into the main script. I can see that this would make it perhaps more organised, and easier to read and edit. Additionally, they recommended adding a 'README' file in order to give more information about what the coding project is about, and perhaps some start up instructions to help them run it more smoothly. I agree that this would be useful, but was under the impression that it was not something we were supposed to do for this exact assignment. Additionally, while renv appeared to work for them, one package was missed (Hmisc), and requried separate installation. They also needed to add the line 'interaction_plot' in order to make it print, though this had not been the case when I ran the chunk myself. Regardless, if it aids in reproducibility, it should be included.

-   *What did you learn about writing code for other people?*
I find writing code for other people (and myself) difficult. However, I think that if you treat your code as if you are writing it for someone else, it will be much more useful to you in the future, as you are unlikely to remember much of the details of it, and so clear explanations and comments help. Likewise, if it can run on someone else's device, then it can run on yours in the future, if features such as 'renv' are used to preserve package versions.

-   Maitner et al. Code sharing increases citations, but remains uncommon. <https://doi.org/10.21203/rs.3.rs-3222221/v1>
-   Trisovic et al. A large-scale study on research code quality and execution. <https://rdcu.be/dZB7x>
-   A Rock-Star Researcher Spun a Web of Lies—and Nearly Got Away with It. <https://thewalrus.ca/a-rock-star-researcher-spun-a-web-of-lies-and-nearly-got-away-with-it/>
-   Gomes et al. Why don't we share data and code? Perceived barriers and benefits to public archiving practices <https://doi.org/10.1098/rspb.2022.1113>
